---
title: "Reproducible-Research"
author: "prem"
date: "July 21, 2017"
output: html_document
---

# 1.Loading and preprocessing the data.
The below mentioed code is used to load the data in to an object.
```{r,echo=TRUE,message=FALSE,warning=FALSE}
act <- read.csv("C:/Users/premsanth/Desktop/Data science coursera/R_programming/repdata_data_activity/activity.csv")
library(lubridate)
act$date <- ymd(act$date) ##This line transform the data type of the date field.
```

# 2.What is mean total number of steps taken per day?

```{r,echo=TRUE,message=FALSE,warning=FALSE}
library(dplyr)
Total_steps_perday <- tapply(X=act$steps,INDEX = act$date,FUN=sum,na.rm=TRUE)
Total_steps_perday_mat<- data.frame(Date = rownames(Total_steps_perday),Total_steps_perday,row.names=NULL) # To convert the array in to dataframe.
```

### The sample of total number of steps taken per day is (only head of the table is printed below)

```{r,echo=TRUE,message=FALSE,warning=FALSE}
print(head(Total_steps_perday_mat))
```

### A histogram of the total number of steps taken each day

```{r,echo=TRUE,message=FALSE,warning=FALSE}
hist(Total_steps_perday)
```

### The mean and median of the total number of steps taken per day.
```{r,echo=TRUE,message=FALSE,warning=FALSE}
(summary(Total_steps_perday))[c("Mean","Median")]
```

# 3.The average daily activity pattern.
### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days.
```{r,echo=TRUE,message=FALSE,warning=FALSE}
library(ggplot2)
library(reshape2)
Average_steps_Interval <- tapply(X=act$steps,INDEX = act$interval,FUN=mean,na.rm=TRUE)
ggplot(melt(Average_steps_Interval),aes(Var1,value))+geom_line()+geom_point()+scale_x_discrete(limits=seq(0,2355,by=150))+labs(x= "5 minitues interval",y="number of steps")
```


#### 5-minute interval contains the maximum number of steps

The 5-minute interval contains the maximum number of steps is `r names(Average_steps_Interval[grep(max(Average_steps_Interval),Average_steps_Interval)])`

# 4.Imputing missing values.
Total number of missing values in the dataset is `length(which(!complete.cases(act)))`
### Filling in all of the missing values in the dataset.
The code for Filling in all of the missing values in the dataset.
```{r,echo=TRUE,message=FALSE,warning=FALSE}
imput_act <- act
imput_act[which(is.na(imput_act$steps)),"steps"] <- as.vector(Average_steps_Interval[(as.character(imput_act[which(is.na(imput_act$steps)),"interval"]))])

```

The head of new dataset that is equal to the original dataset but with the missing data filled in is
```{r,echo=TRUE,message=FALSE,warning=FALSE}
head(imput_act)
```

## For the new dataset.

### total daily number of steps
```{r,echo=TRUE,message=FALSE,warning=FALSE}
Total_steps_perday_new <- tapply(X=imput_act$steps,INDEX = imput_act$date,FUN=sum,na.rm=TRUE)
```

### Mean and median

```{r,echo=TRUE,message=FALSE,warning=FALSE}
(summary(Total_steps_perday_new))[c("Mean","Median")]
```

### histogram

```{r,echo=TRUE,message=FALSE,warning=FALSE}
hist(Total_steps_perday_new)
```

# Activity patterns between weekdays and weekends.
### The code for creating a new factor variable in the dataset.

```{r,echo=TRUE,message=FALSE,warning=FALSE}
imput_act$days <- factor(wday(imput_act$date))
levels(imput_act$days) <- list(week_day=c("2","3","4","5","6"), week_end=c("1","7"))
Average_steps_Interval_weekday <- tapply(X=act$steps,INDEX = list(imput_act$days,imput_act$interval),FUN=mean,na.rm=TRUE)
```

## plot containing a time series plot for weekday and week end.

```{r,echo=TRUE,message=FALSE,warning=FALSE}
ggplot(melt(Average_steps_Interval_weekday),aes(x=Var2,y=value))+facet_grid(Var1~.)+geom_line()+theme_light()+labs(x= "5 minitues interval",y="number of steps")
```

# End of the document.



